### 2022-09-30 cover_letter/coverletter.tex
cover letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Lined Cover Letter
% LaTeX Template
% Version 1.0 (1/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Matthew J. Miller
% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
\makeatother

%\makeatletter
%\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
%\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
%\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
%\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
%\makeatother

\usepackage{url}
% Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options

\usepackage{charter} % Use the Charter font for the document text

\newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
\includegraphics[width=1.2\linewidth]{amu.jpg}}
{\hspace{3.5in}
\includegraphics[width=0.3\linewidth]{CNRS-Logo.png}}
} % Company/institution logo at the top left of the page
\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=10pt}% Slightly decrease the height of the signature field
% Print a phone number under the sender's address
%\newlfmP{addrfromphone} % Print a phone number under the sender's address
\newlfmP{addrfromemail} % Print an email address under the sender's address
%\PhrPhone{Phone} % Customize the "Telephone" text
\PhrEmail{Email} % Customize the "E-mail" text

\lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{\textit{The authors}, Antoine Grimaldi and Laurent Perrinet } % Name

\addrfrom{
\today\\[12pt]
Laurent Perrinet\\
Institute of Neuroscience of la Timone\\
CNRS - Aix Marseille University\\
27, boulevard Jean Moulin\\
13005 Marseille, France
}

%\phonefrom{(000) 111-1111} % Phone number

\emailfrom{laurent.perrinet@univ-amu.fr} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear editor,} % Greeting text
\closeline{Sincerely yours,} % Closing text

%\nameto{Victor Boutin} % Addressee of the letter above the to address

%\addrto{
%Neural Network Editorial Board
%Recruitment Officer \\ % To address
%The Corporation \\
%123 Pleasant Lane \\
%City, State 12345
%}

%----------------------------------------------------------------------------------------

\begin{document}
\begin{newlfm}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
Please find enclosed a manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}. All the authors have been involved with the work, have approved the manuscript and agreed to its submission.

Our manuscript describes a novel method inspired by neuroscience to overcome some challenges faced in computer vision, notably when dealing with large amounts of data. By using heterogeneous delays on different synapses, this novel spiking neuron method is able to detect spiking motifs and we validate" the method on synthetic and then realistic data. Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as thier analog counterparts.

Our  manuscript follows our submission and presentation during the ``NeuroVision'' workshop. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-sourced once the article will be published.

\end{newlfm}
\end{document}